Explore One Variable EDA -- Facebook Pseudo Data
========================================================

***

### Pseudo-Facebook User Data


```{r Pseudo-Facebook User Data}
getwd()
list.files()
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)

```

***

### Histogram of Users' Birthdays

```{r Histogram of Users\' Birthdays}

library(ggplot2)
qplot(x = dob_day, data = pf) + 
  scale_x_continuous(breaks=1:31)
```

***

### Faceting
- facet_wrap(~variable)
       facet_grid(vertical ~ horizontal)

```{r Faceting}
qplot(x = dob_day, data = pf) + 
  scale_x_continuous(breaks=1:31) +
  facet_wrap(~dob_month, ncol = 4)
```

#### Take another look at our plot. 
There's almost 4,000 users in Jan 1 (huge spike), which means lots of people choose the first one in the drop down menus OR protect the privacy then choose the first one.
***

### Friend Count

```{r Friend Count}
qplot(x = friend_count, data = pf)

# optional 2
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram()
```

***

### Limiting the Axes

```{r Limiting the Axes}
qplot(x = friend_count, data = pf, xlim = c(0, 1000))

# Another option:
qplot(x = friend_count, data = pf) +
  scale_x_continuous(limits = c(0, 1000))
```

***

### Adjusting the Bin Width

```{r}
qplot(x = friend_count, data = pf, binwidth = 50) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0,1000,50))
```
- Many users have less than 25 friends, those are probably new.

### Faceting Friend Count
```{r Faceting Friend Count}
# Create a facet the histogram by gender?
qplot(x = friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
facet_wrap(~gender)
```

***
### Omitting NA Values

```{r Omitting NA Values}
# Option 1: should be careful when using na.omit wrapper since it might ignoring some users that are labeled as male/female but do have
#           some NA values in other colomns such as friend_count.
qplot(x = friend_count, data = na.omit(pf), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
facet_wrap(~gender)

# Option 2:
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
facet_wrap(~gender)

```

***

### Statistics 'by' Gender
- Who has more friends? Male or Female? 

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)

```
***

### Tenure
- the 'color' determines the color outline of objects in a plot
- the 'fill' determines the color of the area inside objects in a plot

```{r Tenure}
qplot(x = tenure, data = pf,
      color = I('black'), fill = I('#099DD9'))
```
```{r}
# measure tenure by years rather than days
qplot(x = tenure/365, data = pf, binwidth = .25,
      color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1,7,1), limits = c(0,7))
```

***

### Labeling Plots

```{r Labeling Plots}
library(ggplot2)
qplot(x = tenure/365, data = pf,
      xlab = 'Number of years using Facebook',
      ylab = 'Number of users in sample',
      color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1,7,1), limits = c(0,7))
```

***

### User Ages

```{r User Ages}
qplot(x = age, data = pf, binwidth = 1,
      xlab = 'Age',
      ylab = 'Number of Users',
      color = I('black'), fill = I('#5760AB'))
```

***

### Transforming Data

- engagement variables -- all have very long tails. In statistics, the data is over dispersed. Thus, we need to transform these values to short the tails.
```{r}
qplot(x = friend_count, data = pf)

summary(pf$friend_count)

# transform the variable by taking the log (natural log/ log base2/log base10)

summary(log10(pf$friend_count + 1))  # add 1 to avoid 0 to be negative infinity in log

# # transform the variable by taking the sqrt
summary(sqrt(pf$friend_count))

```
```{r}
# create multiple plots on one page by using scales
library(gridExtra)

# Option 1: 
p1 <- qplot(x = friend_count, data = (pf))
p2 <- qplot(x = log10(pf$friend_count + 1), data = pf)
p3 <- qplot(x = sqrt(pf$friend_count), data = pf)
grid.arrange(p1,p2,p3, ncol=1)

# Option 2:
p1 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()
grid.arrange(p1,p2,p3, ncol = 1)


```
***

### Add a Scaling Layer


```{r Add a Scaling Layer}
# Two ways to transfer the variable: 

# 1. used a wrapper right around the variable

logScale <- qplot(x = log10(friend_count), data = pf)

# 2. used a scaling layer
countScale <- ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_log10()

grid.arrange(logScale, countScale, ncol = 2)

# The difference between those two plots: 
# -- using the log10 wrapper will label the x axis in log units where the layer will reflect the real numbers
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_log10()

# OR can do it in this way
qplot(x = friend_count, data = pf) +
  scale_x_log10()
```

***

### Frequency Polygons
- frequency polygons are similar to histograms, but draw a curve connecting the counts in a histogram. 
- overlay these histograms together.

```{r Frequency Polygons}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50)) +
facet_wrap(~gender)

qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50))


qplot(x = friend_count, y = ..count../sum(..count..), 
      data = subset(pf, !is.na(gender)), 
      xlab = 'Friends Count',
      ylab = 'Proportion of Users with that friend count',
      binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50))

```

***

### Likes on the Web

```{r Likes on the Web}

qplot(x = www_likes, data = subset(pf, !is.na(gender)),
      geom = 'freqpoly', color = gender) +
  scale_x_continuous()

# deal with long tail with log10 transform
qplot(x = www_likes, data = subset(pf, !is.na(gender)),
      geom = 'freqpoly', color = gender) +
  scale_x_continuous() +
  scale_x_log10()

# Question: What's the www_like count for males? Which gender has more www_likes?

by(pf$www_likes, pf$gender, sum)  #  1430175 (male) < 3507665 (female)

```

***

### Box Plots
- another type of visualization for seeing the distribution of a variable
- the female's median line seems slightly higher than males

```{r Box Plots}

qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot')
# the horizontal line is the median for the two box plots
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot', ylim = c(0, 1000)) # Adjust the code to focus on users who have friend counts between 0 and 1000
```

***

### Box Plots, Quartiles, and Friendships

```{r Box Plots, Quartiles, and Friendships}
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot') + 
  coord_cartesian(ylim = c(0, 250))

by(pf$friend_count, pf$gender, summary)

```

#### On average, who initiated more friendships in our sample: men or women?

```{r Friend Requests by Gender}
qplot(x = gender, y = friendships_initiated,
      data = subset(pf, !is.na(gender)), geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 500))

# slightly higher them zoom in again
qplot(x = gender, y = friendships_initiated,
      data = subset(pf, !is.na(gender)), geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 150))

by(pf$friendships_initiated, pf$gender, summary)

```

***

### Getting Logical
- to a new binary variable that has only true or false

```{r Getting Logical}
summary(pf$mobile_likes)
summary(pf$mobile_likes > 0)
```
What percent of check in using mobile?
```{r}
mobile_check_in <- NA # create a new variable in the data frame with NA values
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
sum(pf$mobile_check_in == 1) / length(pf$mobile_check_in)
```
***
