Lesson 4
======================================================

### Scatterplots

```{r Scatterplots}

library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

qplot(x = age, y = friend_count, data = pf)


```

***

### ggplot Syntax
- ggplot function is similar to qplot with slightly different syntax

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_point() +
  xlim(13,90)
# set the age range from 30 - 90 more reasonable

summary(pf$age)

```

***

### Overplotting
- Some of these points are spread out from one another, while others are stacked right on top of each other. 
- set the transparency of the points using the alpha parameter and geom point

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20) +
  xlim(13,90) 

# we can see the bulk of young users really have friend counts below 1000

```

***
#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13,90) +
  coord_trans(y = "sqrt")

```
***
#### Examine the relationship between friendships_initiated (y) and age (x)
```{r}
ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
  geom_point(alpha = 1/10, position = position_jitter(h = 0)) +
  xlim(10,100) +
  coord_trans(y = "sqrt")
```

***
### Overplotting and Domain Knowledge
- Transform the axes, using the percentage of their friend count, which show the truth for some have fewer friends. Compare the percentage rather than amount. 

***

### Conditional Means
- dplyr package: split up a data frame and apply a function to some parts of the data. Such as: filter(), group_by(), mutate(), arrange().

```{r Conditional Means}
library(dplyr)
age_groups <- group_by(pf, age)

pf.fc_by_age <- summarise(age_groups,
                           friend_count_mean = mean(friend_count),
                           friend_count_median =median(as.numeric(friend_count)),
                           n = n())

head(pf.fc_by_age,10)

# conditional means alternate code
library(dplyr)

pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(as.numeric(friend_count)),
            n = n()) %>%
  arrange(age)

head(pf.fc_by_age, 20)
```

```{r Conditional Means Plot}
# Plot mean friend count vs. age using a line graph.
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + 
  geom_line() + xlim(10,100)
```

***

### Overlaying Summaries with Raw Data

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  coord_cartesian(xlim = c(13,70), ylim = c(0, 1000)) +
  geom_point(alpha = 0.05,
             position = position_jitter(h = 0),
             color = 'orange') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', 
            fun.y = quantile, fun.args = list(probs = .9),
            linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', 
            fun.y = quantile, fun.args = list(probs = .5),
            linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', 
            fun.y = quantile, fun.args = list(probs = .1),
            linetype = 2, color = 'blue') 

# set 10%, 50% and 90% quantile on plot
```
***
### Correlation

```{r Correlation}
cor.test(pf$age, pf$friend_count, method = 'pearson')

# Another option:
with(pf, cor.test(age, friend_count, method = 'pearson'))
```
***

### Correlation on Subsets
- Based on the plot above, it seems that the relationship between age and friend count is not linear. It isn't monotonic, either increasing or decreasing.
- We maybe don't want to include the older ages, which likely is incorrect.

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***
### Correlation Methods
- monotonic relationsip, such as a rank correlation measures like Spearman.
```{r Monotonic Correlation}
with(subset(pf, age <= 70), cor.test(age, friend_count, method = 'spearman'))
```
***

## Create Scatterplots
- Create a scatterplot of likes_received (y) vs. www_likes_received (x).

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point() 
```

***

### Strong Correlations
- from the plot, there're some funky outliers here. To determine good x and y limits for axis, we can look at 95th percentile

```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point() + 
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'red')
```

- What's the correlation betwen the two variables? 

```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received)
```

***

### Moira on Strong Correlation FB Case
- FB data, a lot of data that I worked with is correlated with each other. For example, "how many status updated someone posts in a month"is usually highly correlated with "how many days in last month they logged in" OR " how many friends they have" OR "how many photos they uploaded in last month". All of these variables are typically very highly related because they all kind of measure the same thing. It's how engaed someone is.

- Regression : variables are independent with each other.

***

```{r Temp vs Month}
library(ggplot2)
library(alr3)

data(Mitchell)
?Mitchell

# Create a scatterplot of temperature (Temp) vs. months (Month)
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()
cor.test(Mitchell$Month, Mitchell$Temp)

```

***

### Making Sense of Data
- To add a layer to make the months to be discrete

```{r Making Sense of Data}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point() +
  scale_x_continuous(breaks = seq(0,203,12))

range(Mitchell$Month) # to determine the month upper bound

```

***

### A New Perspective
- When stretch out of the graph, it's noticed that a cyclical pattern. It's almost like a sin or cosin graph.And this makes sense with the seasons in Nebraska; the fluctuation in the tempeturure every 12 months. 
- Make your data on the context!

***

### Understanding Noise: Age to Age Months
- 

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x = age, y = friend_count_mean), 
       data = subset(pf.fc_by_age, age < 71)) +
  geom_line()

head(pf.fc_by_age, 10)

pf.fc_by_age[17:19,]
```

***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- pf$age + (12 - pf$dob_month) / 12
```

Programming Assignment
```{r Programming Assignment}
library(dplyr)
age_with_months_groups <- group_by(pf, age_with_months)

pf.fc_by_age_months2 <- summarise(age_with_months_groups,
                           friend_count_mean = mean(friend_count),
                           friend_count_median =median(as.numeric(friend_count)),
                           n = n())

head(pf.fc_by_age_months2,10)


```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_months2, age_with_months < 71)) +
  geom_line()
# here is the much nosier plot 
```

***

### Smoothing Conditional Means
- LOESS and LOWESS ( locally weighted scatterplot smoothing) are two methods that combine multiple regression in a k-nearest-neighbor-based meta-model. 


```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_months2, age_with_months < 71)) +
  geom_line() +
  geom_smooth() 

p2 <- ggplot(aes(x = age, y = friend_count_mean), 
       data = subset(pf.fc_by_age, age < 71)) +
  geom_line() +
  geom_smooth()


library(gridExtra)
grid.arrange(p1, p2, ncol = 1)
  
```
### Which Plot to Choose?

- by decreasing the size and bins and increasing the number of bins, we can see the noise is much worse in the age_with_months plot
- On the other hand, we could go the other direction and increase the size of the bins. Say, we could lump everyone together whose age falls under a multiple of five, in plot 3. 

```{r Chose the plot}

p1 <- ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_months2, age_with_months < 71)) +
  geom_line()

p2 <- ggplot(aes(x = age, y = friend_count_mean), 
       data = subset(pf.fc_by_age, age < 71)) +
  geom_line()
  
p3 <- ggplot(aes(x = round(age / 5) * 5, y = friend_count_mean), 
       data = subset(pf.fc_by_age, age < 71)) +
  geom_line(stat = 'summary', fun.y = mean)

library(gridExtra)
grid.arrange(p1, p2, p3, ncol = 1)
```

***



