Lesson 5
========================================================
***

### Third Qualitative Variable

```{r Third Qualitative Variable}
library(ggplot2)
# we have noticed that female users have more friends on average than male users, is that just because female users have a different age distribution? Or maybe conditional on age?

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)
# add mean to each gender to the box plots, using stat summary

# Male users are younger,we might actually think a simple male to female comparision doesn't capture their substantial differences in friend count. So, let's look at median friend count by age and gender instead.
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)

```

***

### Plotting Conditional Summaries
- the codes different but the plot is the same result

```{r Plotting Conditional Summaries}
# Create a new data frame ('pf.fc_by_age_gender'), that contains
# information on each age AND gender group.

library(dplyr)
age_with_gender_groups <- group_by(pf, age, gender)


pf.fc_by_age_gender <- summarise(age_with_gender_groups,
                           mean_friend_count = mean(friend_count),
                           median_freind_count = median(as.numeric(friend_count)),
                           n = n())

head(pf.fc_by_age_gender)

# Better method:
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(as.numeric(friend_count)),
                           n = n()) %>%
  ungroup() %>%
  arrange(age)

head(pf.fc_by_age_gender)

ggplot(aes(x = age, y = median_friend_count),
       data = pf.fc_by_age_gender) + 
  geom_line(aes(color = gender))

# From the plot, it seems like that the gender difference is largest for young users.
```

***

### Reshaping Data
- Question: How many times more friends does the average female user have than the male user?
- To answer that question, we need to rearrange the data, from a long format to a wide format. one colomn (gender) -- two colomns (female and male)

```{r Reshape package}

install.packages('reshape2')
library(reshape2)
pf.fc_by_age_gender_wide <- dcast(pf.fc_by_age_gender, # the dataset
                                  age ~ gender, # left list the variable want to keep, right side is the variable need to split
                                  value.var = 'median_friend_count')
head(pf.fc_by_age_gender_wide)

```

```{r Tidyr Package}
install.packages('tidyr')
library(tidyr)
spread(subset(pf.fc_by_age_gender, 
       select = c('gender', 'age', 'median_friend_count')), 
       gender, median_friend_count)

```

***

### Ratio Plot
- Plot the ratio of the female to male median friend counts 

```{r Ratio Plot}
library(ggplot2)
ggplot(aes(x = age, y = female),
       data - pf.fc_by_age_gender_wide) +
  geom_line() +
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)

```

***

### Third Quantitative Variable

```{r Third Quantitative Variable}
# floor() round down to the nearest interger. ceiling() round down to the cloest year
pf$year_joined <- floor(2014 - pf$tenure/365) 
head(pf) 
```

***

### Cut a Variable

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)

# create 4 bins from 2004-2009, 2009-2011,2011-2012,2012-2014
pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together

```{r Plotting it All Together}
table(pf$year_joined.bucket, useNA = 'ifany')
ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)

# create a line graph of friend_count vs. age, so that each year_joined.bucket is a line tracking the median user friend_count across age
ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary', fun.y = median)

# From the plot above, we can see users with a longer tenure tend to have higher friend counts, with the exception of the older users

```

***

### Plot the Grand Mean

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)

# ploting a grand mean is a good remindar that much of the data in the sample is about members of recent cohorts
```

***

### Friending Rate
- check the freind count each day for each user 

```{r Friending Rate}

with(subset(pf, tenure > 1), summary(friend_count / tenure))

# the max is 417, which is an outlier since the 3rd Qu is only 0.56
```

***

```{r Friendships Initiated}
# create a line graph of mean of friendships_initiated per day vs. tenure colored by year_joined.bucket
# also need to subset the data to only consider user with at least one day of tenure
ggplot(aes(x = tenure, y = friendships_initiated / tenure), 
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary', fun.y = mean)

# it appears that users with more tenure typically initiate less friendships

```

***

### Bias-Variance Tradeoff Revisited
- there's a lot of noise in the graph since we're plotting the mean of y for every possible tenure x value
- we can adjust this noise by bending x-axis differently

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)
# use geom_smooth to automatically adjust 
ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket)) 

# we still see that the friendships initiated declines as tenure increases
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

