RED WINE QUALITY EDA -- Freena
========================================================
# 1. Introduction

In this project, I explore the quality of red wine. This data set contains 1,599 red wine records with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent). (cite1)

Through these variables, I'd like to uncover which chemical properties influence the quality of red wines and Which variable affect most? Perhaps the results can help me in choosing red wine next time at the Wine shop. 

## 1.1 Preparations
```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code chunk.
library(ggplot2)
library(GGally)
library(gridExtra)
library(dplyr)
```

## 1.2 Intro to The Dataset 
```{r echo=FALSE, Intro_to_the_Data}
# Load the Data
wine <- read.csv('wineQuality.csv')

# check the first 10 records and basic data structure
head(wine, 10)
str(wine)

```

##  1.3 Description of Variables:

For the structure of dataset, we have 11 variables, in which the 'quality' is the dependent variable and other 10 properties are the independent variables. In addition, there's no missing values in these variables and the dataset itself is pretty tidy.

1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (tartaric acid - g / dm^3)
2 - volatile acidity: the amount of acetic acid in wine (acetic acid - g / dm^3)
3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines (g / dm^3)
4 - residual sugar: the amount of sugar remaining after fermentation stops (g / dm^3)
5 - chlorides: the amount of salt in the wine (sodium chloride - g / dm^3)
6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (mg / dm^3)
7 - total sulfur dioxide: amount of free and bound forms of S02 (mg / dm^3)
8 - density: the density of water is close to that of water depending on the percent alcohol and sugar content (g / cm^3)
9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale
10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels (potassium sulphate - g / dm3)
11 - alcohol: the percent alcohol content of the wine (% by volume)
12 - quality (score between 0 and 10)

##  1.4 Data Wrangling and Preprocessing:

Take a quick look of the dataset, I can find some small issues need to fix before the EDA. Even though this dataset is pretty tidy, fixing these issues will make the dataset perfect to conduct further exploration.

1. Change the variable names to make all variable names identical -- change 'citric.acid' to 'citric.acidity'.
2. Standardize the measure method of the chemical properties of the wine -- change mg / dm^3 to g / dm^3 for 'total sulfur dioxide' and 'free sulfur dioxide'.
3. drop the 'X' variable since it's simply a index that duplicate with the default index.

```{r Data Wrangling}

# 1. rename col
colnames(wine)[colnames(wine)=="citric.acid"] <- "citric.acidity"

# 2. change measurement 
wine$total.sulfur.dioxide <- wine$total.sulfur.dioxide / 1000
wine$free.sulfur.dioxide <- wine$free.sulfur.dioxide / 1000

# 3. drop X
drops <- c('X')
wine <- wine[, !names(wine) %in% drops]

# 4. check the result
tail(wine)
```


# 2. Univariate Plots Section

## 2.1 Dependent Variable -- Quality

From the structure of the dataset, we can see that all the variables are numerics except the quality. It's ok to keep this data type without changing it from integer to factor since I need to create a new variable called 'quality.bucket' to set three levels of wine quality: low/median/high based on the true distribution of quality (from 1 to 8).

```{r echo=FALSE, Distribution of Quality}

# check the distribution of quality
summary(wine$quality)

p1 <- ggplot(aes(x = quality), data = wine) +
  geom_bar(fill = '#9999CC')

# cut the quality.bucket -- 1-4 low/ 5-6 medium/ 7-8 high 
wine$quality.bucket <- cut(wine$quality, c(1,4,6,8), labels = c('Low','Medium', 'High'))
summary(wine$quality.bucket)

p2 <- ggplot(aes(x = quality.bucket), data = wine) +
  geom_bar(fill = '#9999CC')

grid.arrange(p1,p2)

```

## 2.2 Independnet Variables 

For all the independent variables, I check all the statsitical summarys and plot the histograms of to check their distribution as the same time visualize the mean (red line) and median (green line) in the plots.

```{r echo=FALSE, Distribution of Independent Variables}

summary(wine)
geom_histogram()

q1 <- ggplot(aes(x = fixed.acidity), data = wine) + geom_bar() + geom_vline(aes(xintercept = mean(fixed.acidity)), color = 'red') + geom_vline(aes(xintercept = median(fixed.acidity)), color = 'green')

q2 <- ggplot(aes(x = volatile.acidity), data = wine) + geom_bar() + geom_vline(aes(xintercept = mean(volatile.acidity)), color = 'red') + geom_vline(aes(xintercept = median(volatile.acidity)), color = 'green')

q3 <- ggplot(aes(x = citric.acidity), data = wine) + geom_bar() + geom_vline(aes(xintercept = mean(citric.acidity)), color = 'red') + geom_vline(aes(xintercept = median(citric.acidity)), color = 'green')

q4 <- ggplot(aes(x = residual.sugar), data = wine) + geom_bar() + geom_vline(aes(xintercept = mean(residual.sugar)), color = 'red') + geom_vline(aes(xintercept = median(residual.sugar)), color = 'green')

q5 <- ggplot(aes(x = chlorides), data = wine) + geom_bar() + geom_vline(aes(xintercept = mean(chlorides)), color = 'red') + geom_vline(aes(xintercept = median(chlorides)), color = 'green')

q6 <- ggplot(aes(x = free.sulfur.dioxide), data = wine) + geom_bar() + geom_vline(aes(xintercept = mean(free.sulfur.dioxide)), color = 'red') + geom_vline(aes(xintercept = median(free.sulfur.dioxide)), color = 'green')

q7 <- ggplot(aes(x = total.sulfur.dioxide), data = wine) + geom_bar() + geom_vline(aes(xintercept = mean(total.sulfur.dioxide)), color = 'red') + geom_vline(aes(xintercept = median(total.sulfur.dioxide)), color = 'green')

q8 <- ggplot(aes(x = density), data = wine) + geom_bar() + geom_vline(aes(xintercept = mean(density)), color = 'red') + geom_vline(aes(xintercept = median(density)), color = 'green')

q9 <- ggplot(aes(x = pH), data = wine) + geom_bar() + geom_vline(aes(xintercept = mean(pH)), color = 'red') + geom_vline(aes(xintercept = median(pH)), color = 'green')

q10 <- ggplot(aes(x = sulphates), data = wine) + geom_bar() + geom_vline(aes(xintercept = mean(sulphates)), color = 'red') + geom_vline(aes(xintercept = median(sulphates)), color = 'green')

q11 <- ggplot(aes(x = alcohol), data = wine) + geom_bar() + geom_vline(aes(xintercept = mean(alcohol)), color = 'red') + geom_vline(aes(xintercept = median(alcohol)), color = 'green')


grid.arrange(q1,q2,q3,q4,q5,q6,q7,q8,q9,q10,q11,ncol = 3)

```

From the plots and summarys above, we can see that distributions of the variables are varied. Many variables are over-dispered with a very long tails. Thus, I need to transform these values to short the tails. Dependent on different distributions, I transform the variables ('residual.sugar', 'chlorides', 'total.sulfur.dioxide', 'sulphates') by taking the log base10 and transform the variables (citric.acidity', 'free.sulfur.dioxide') by taking the sqrt. 

I also apply the transformation to the variables ('fixed.acidity', 'volatile.acidity', 'citric.acidity') but there're no significant change in the distribution, for that reason, I keep the original distribution status for these three variables. 

Besides, no need to change in the distributions of three variables: 'density','pH', and 'alcohol'.

In short, I trim the long tails for most variables by taking log base10 and sqrt at the same time adjust the binwidth of them for better visualization. 

```{r echo=FALSE, Transforming Data}
q1 <- ggplot(aes(x = fixed.acidity), data = wine) + geom_histogram(binwidth = 1) + geom_vline(aes(xintercept = mean(fixed.acidity)), color = 'red') + geom_vline(aes(xintercept = median(fixed.acidity)), color = 'green')

q2 <- ggplot(aes(x = volatile.acidity), data = wine) + geom_histogram(binwidth = 0.1) + geom_vline(aes(xintercept = mean(volatile.acidity)), color = 'red') + geom_vline(aes(xintercept = median(volatile.acidity)), color = 'green') 

q3 <- ggplot(aes(x = citric.acidity), data = wine) + geom_histogram(binwidth = 0.05) + geom_vline(aes(xintercept = mean(citric.acidity)), color = 'red') + geom_vline(aes(xintercept = median(citric.acidity)), color = 'green') + scale_x_sqrt()

q4 <- ggplot(aes(x = residual.sugar), data = wine) + geom_histogram(binwidth = 0.05) + geom_vline(aes(xintercept = mean(residual.sugar)), color = 'red') + geom_vline(aes(xintercept = median(residual.sugar)), color = 'green') +  scale_x_log10()

q5 <- ggplot(aes(x = chlorides), data = wine) + geom_histogram(binwidth = 0.05) + geom_vline(aes(xintercept = mean(chlorides)), color = 'red') + geom_vline(aes(xintercept = median(chlorides)), color = 'green') +  scale_x_log10()


q6 <- ggplot(aes(x = free.sulfur.dioxide), data = wine) + geom_histogram(binwidth = 0.01) + geom_vline(aes(xintercept = mean(free.sulfur.dioxide)), color = 'red') + geom_vline(aes(xintercept = median(free.sulfur.dioxide)), color = 'green') +  scale_x_sqrt()

q7 <- ggplot(aes(x = total.sulfur.dioxide), data = wine) + geom_histogram(binwidth = 0.1) + geom_vline(aes(xintercept = mean(total.sulfur.dioxide)), color = 'red') + geom_vline(aes(xintercept = median(total.sulfur.dioxide)), color = 'green') + scale_x_log10()

q8 <- ggplot(aes(x = density), data = wine) + geom_histogram(binwidth = 0.001) + geom_vline(aes(xintercept = mean(density)), color = 'red') + geom_vline(aes(xintercept = median(density)), color = 'green')

q9 <- ggplot(aes(x = pH), data = wine) + geom_histogram(binwidth = 0.1) + geom_vline(aes(xintercept = mean(pH)), color = 'red') + geom_vline(aes(xintercept = median(pH)), color = 'green')

q10 <- ggplot(aes(x = sulphates), data = wine) + geom_histogram(binwidth = 0.05) + geom_vline(aes(xintercept = mean(sulphates)), color = 'red') + geom_vline(aes(xintercept = median(sulphates)), color = 'green') + scale_x_log10()

q11 <- ggplot(aes(x = alcohol), data = wine) + geom_histogram(binwidth = 0.5) + geom_vline(aes(xintercept = mean(alcohol)), color = 'red') + geom_vline(aes(xintercept = median(alcohol)), color = 'green')

grid.arrange(q1,q2,q3,q4,q5,q6,q7,q8,q9,q10,q11, ncol = 3)


```

# Univariate Analysis

### What is the structure of your dataset?

For the structure of dataset, we have 11 variables, in which the 'quality' is the dependent variable and other 10 properties are the independent variables. In addition, there's no missing values in these variables and the dataset itself is pretty tidy.

### What is/are the main feature(s) of interest in your dataset?

The 'quality' is the dependent variable which I am interest most. I also feel interest in alcohol, residual.sugar, density, and pH, as those might be the important factors to red wine quality rating.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
citric.acidity and chlorides

### Did you create any new variables from existing variables in the dataset?

Yes. I create a new variable called 'quality.bucket' to set three levels of wine quality: low/median/high based on the true distribution of quality (from 1 to 8).

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

Many variables are over-dispered with a very long tails. Thus, I need to transform these values to short the tails. Dependent on different distributions, I transform the variables ('residual.sugar', 'chlorides', 'total.sulfur.dioxide', 'sulphates') by taking the log base10 and transform the variables (citric.acidity', 'free.sulfur.dioxide') by taking the sqrt. 


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

Due to privacy and logistic issues, only physicochemical (inputs) and sensory (the output) variables are available (e.g. there is no data about grape types, wine brand, wine selling price, etc.).

These datasets can be viewed as classification or regression tasks.

The classes are ordered and not balanced (e.g. there are munch more normal wines than excellent or poor ones). Outlier detection algorithms could be used to detect the few excellent or poor wines. Also, we are not sure if all input variables are relevant.


> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!